---
title: "IRT without the normality assumption"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{IRT without the normality assumption}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
fontsize: 12pt
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(IRTest)
library(ggplot2)
library(gridExtra)
```

# 0. Introduction 

+ **IRTest** is a useful tool for $\mathcal{\color{red}{IRT}}$ (item response theory) parameter $\mathcal{\color{red}{est}}$imation, especially when the violation of normality assumption on latent distribution is suspected.

+ **IRTest** deals with uni-dimensional latent variable.

+ For missing values, **IRTest** adopts full information maximum likelihood (FIML) approach.

+ In **IRTest**, including the conventional usage of Gaussian distribution, several methods are available for estimation of latent distribution:
    + empirical histogram method,
    + two-component Gaussian mixture distribution,
    + Davidian curve,
    + kernel density estimation,    
    + log-linear smoothing.

## Installation

The CRAN version of **IRTest** can be installed on R-console with:

```
install.packages("IRTest")
```

For the development version, it can be installed on R-console with:
```
devtools::install_github("SeewooLi/IRTest")
```

## Functions

Followings are the functions of **IRTest**.

  + `IRTest_Dich` is the estimation function when all items are *dichotomously* scored.
  
  + `IRTest_Poly` is the estimation function when all items are *polytomously* scored.
  
  + `IRTest_Cont` is the estimation function when all items are *continuously* scored.
  
  + `IRTest_Mix` is the estimation function for *a mixed-format test*, a test comprising both dichotomous item(s) and polytomous item(s).
  
  + `factor_score` estimates factor scores of examinees.
  
  + `coef_se` returns standard errors of item parameter estimates.
  
  + `best_model` selects the best model using an evaluation criterion.
  
  + `item_fit` tests the statistical fit of all items individually.
  
  + `inform_f_item` calculates the information value(s) of an item.

  + `inform_f_test` calculates the information value(s) of a test.
    
  + `plot_item` draws item response function(s) of an item.
  
  + `reliability` calculates marginal reliability coefficient of IRT.
  
  + `latent_distribution` returns evaluated PDF value(s) of an estimated latent distribution.

  + `DataGeneration` generates several objects that can be useful for computer simulation studies. Among these are simulated item parameters, ability parameters and the corresponding item-response data.
  
  + `dist2` is a probability density function of two-component Gaussian mixture distribution.
  
  + `original_par_2GM` converts re-parameterized parameters of two-component Gaussian mixture distribution into original parameters.  
  
  + `cat_clps` recommends category collapsing based on item parameters (or, equivalently, item response functions).
  
  + `recategorize` implements the category collapsing.
  
  + For S3 methods, `anova`, `coef`, `logLik`, `plot`, `print`, and `summary` are available.


\

---

\break

# 1. Dichotomous items

+ **Preparing data**

The function `DataGeneration` can be used in a preparation step.
This function returns a set of artificial data and the true parameters underlying the data.

```{r}
Alldata <- DataGeneration(model_D = 2,
                          N=1000,
                          nitem_D = 15,
                          latent_dist = "2NM",
                          d = 1.664,
                          sd_ratio = 2,
                          prob = 0.3)

data <- Alldata$data_D
theta <- Alldata$theta
colnames(data) <- paste0("item", 1:15)
```

\

+ **Analysis** (parameter estimation)

```{r, results='hide', message=FALSE}
Mod1 <- IRTest_Dich(data = data,
                    model = 2,
                    latent_dist = "LLS",
                    h=4)
```

\

+ **Results**

```{r, message=FALSE, fig.align='center', fig.height=6, fig.width=6}
### Summary
summary(Mod1)

### Log-likelihood
logLik(Mod1)

### The estimated item parameters
coef(Mod1)

### Standard errors of the item parameter estimates
coef_se(Mod1)

### The estimated ability parameters
fscore <- factor_score(Mod1, ability_method = "MLE")
plot(theta, fscore$theta)
abline(b=1, a=0)

### Standard errors of ability parameter estimates
plot(fscore$theta, fscore$theta_se)
```


* The result of latent distribution estimation

```{r, fig.align='center', fig.asp=.6, fig.width=6}
plot(Mod1, mapping = aes(colour="Estimated"), linewidth = 1) +
  lims(y = c(0, .75))+
  geom_line(
    mapping=aes(
      x=seq(-6,6,length=121), 
      y=dist2(seq(-6,6,length=121), prob = .3, d = 1.664, sd_ratio = 2), 
      colour="True"),
    linewidth = 1)+
  labs(title="The estimated latent density using '2NM'", colour= "Type")+
  theme_bw()
```

* Item response function

```{r, fig.align='center', fig.asp=0.8, fig.width=6}
p1 <- plot_item(Mod1,1)
p2 <- plot_item(Mod1,4)
p3 <- plot_item(Mod1,8)
p4 <- plot_item(Mod1,10)
grid.arrange(p1, p2, p3, p4, ncol=2, nrow=2)
```


+ **Item fit**
```{r}
item_fit(Mod1)
```

* Reliability

```{r}
reliability(Mod1)
```

* Posterior distributions for the examinees

Each examinee's posterior distribution is identified in the E-step of the estimation algorithm (i.e., EM algorithm).
Posterior distributions can be found in `Mod1$Pk`.

```{r, fig.align='center', fig.asp=.7, fig.width=6}
set.seed(1)
selected_examinees <- sample(1:1000,6)
post_sample <- 
  data.frame(
    X = rep(seq(-6,6, length.out=121),6), 
    prior = rep(Mod1$Ak/(Mod1$quad[2]-Mod1$quad[1]), 6),
    posterior = 10*c(t(Mod1$Pk[selected_examinees,])), 
    ID = rep(paste("examinee", selected_examinees), each=121)
    )

ggplot(data=post_sample, mapping=aes(x=X))+
  geom_line(mapping=aes(y=posterior, group=ID, color='Posterior'))+
  geom_line(mapping=aes(y=prior, group=ID, color='Prior'))+
  labs(title="Posterior densities for selected examinees", x=expression(theta), y='density')+
  facet_wrap(~ID, ncol=2)+
  theme_bw()
```

* Test information function

```{r, fig.align='center', fig.asp=.8, fig.width=6}
ggplot()+
  stat_function(
    fun = inform_f_test,
    args = list(Mod1)
  )+ 
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 1),
    mapping = aes(color="Item 1")
  )+
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 2),
    mapping = aes(color="Item 2")
  )+
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 3),
    mapping = aes(color="Item 3")
  )+
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 4),
    mapping = aes(color="Item 4")
  )+
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 5),
    mapping = aes(color="Item 5")
  )+
  lims(x=c(-6,6))+
  labs(title="Test information function", x=expression(theta), y='information')+
  theme_bw()
```


----------

\break

# 2. Polytomous items

+ **Preparing data**


```{r}
Alldata <- DataGeneration(model_P = "GRM",
                          categ = rep(c(3,7), each = 7),
                          N=1000,
                          nitem_P = 14,
                          latent_dist = "2NM",
                          d = 1.664,
                          sd_ratio = 2,
                          prob = 0.3)

data <- Alldata$data_P
theta <- Alldata$theta
colnames(data) <- paste0("item", 1:14)
```

\

+ **Analysis** (parameter estimation)

```{r, results='hide', message=FALSE}
Mod1 <- IRTest_Poly(data = data,
                    model = "GRM",
                    latent_dist = "KDE")
```

\

+ **Results**

```{r, message=FALSE, fig.align='center', fig.height=6, fig.width=6}
### Summary
summary(Mod1)

### Log-likelihood
logLik(Mod1)

### The estimated item parameters
coef(Mod1)

### Standard errors of the item parameter estimates
coef_se(Mod1)

### The estimated ability parameters
fscore <- factor_score(Mod1, ability_method = "MLE")
plot(theta, fscore$theta)
abline(b=1, a=0)

### Standard errors of ability parameter estimates
plot(fscore$theta, fscore$theta_se)
```

* The result of latent distribution estimation

```{r, fig.align='center', fig.asp=.6, fig.width=6}
plot(Mod1, mapping = aes(colour="Estimated"), linewidth = 1) +
  stat_function(
    fun = dist2,
    args = list(prob = .3, d = 1.664, sd_ratio = 2),
    mapping = aes(colour = "True"),
    linewidth = 1) +
  lims(y = c(0, .75)) + 
  labs(title="The estimated latent density using '2NM'", colour= "Type")+
  theme_bw()
```

* Item response function

```{r, fig.align='center', fig.asp=0.8, fig.width=6}
p1 <- plot_item(Mod1,1)
p2 <- plot_item(Mod1,4)
p3 <- plot_item(Mod1,8)
p4 <- plot_item(Mod1,10)
grid.arrange(p1, p2, p3, p4, ncol=2, nrow=2)
```


+ **Item fit**
```{r}
item_fit(Mod1)
```

* Reliability

```{r}
reliability(Mod1)
```

* Posterior distributions for the examinees

Each examinee's posterior distribution is identified in the E-step of the estimation algorithm (i.e., EM algorithm).
Posterior distributions can be found in `Mod1$Pk`.

```{r, fig.align='center', fig.asp=.7, fig.width=6}
set.seed(1)
selected_examinees <- sample(1:1000,6)
post_sample <- 
  data.frame(
    X = rep(seq(-6,6, length.out=121),6), 
    prior = rep(Mod1$Ak/(Mod1$quad[2]-Mod1$quad[1]), 6),
    posterior = 10*c(t(Mod1$Pk[selected_examinees,])), 
    ID = rep(paste("examinee", selected_examinees), each=121)
    )

ggplot(data=post_sample, mapping=aes(x=X))+
  geom_line(mapping=aes(y=posterior, group=ID, color='Posterior'))+
  geom_line(mapping=aes(y=prior, group=ID, color='Prior'))+
  labs(title="Posterior densities for selected examinees", x=expression(theta), y='density')+
  facet_wrap(~ID, ncol=2)+
  theme_bw()
```

* Test information function

```{r, fig.align='center', fig.asp=.8, fig.width=6}
ggplot()+
  stat_function(
    fun = inform_f_test,
    args = list(Mod1)
  )+ 
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 1),
    mapping = aes(color="Item 1 (3 cats)")
  )+
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 2),
    mapping = aes(color="Item 2 (3 cats)")
  )+
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 3),
    mapping = aes(color="Item 3 (3 cats)")
  )+
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 8),
    mapping = aes(color="Item 8 (7 cats)")
  )+
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 9),
    mapping = aes(color="Item 9 (7 cats)")
  )+
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 10, "p"),
    mapping = aes(color="Item10 (7 cats)")
  )+
  lims(x=c(-6,6))+
  labs(title="Test information function", x=expression(theta), y='information')+
  theme_bw()
```


----------

\break

\

# 3. Continuous items

+ **Statistical details of the continuous IRT**


<details>
<summary><b>Beta distribution (click)</b></summary>

$$
\begin{align}
f(x) &= \frac{1}{Beta(\alpha, \beta)}x^{\alpha-1}(1-x)^{(\beta-1)} \\
&= \frac{\Gamma(\alpha+\beta)}{\Gamma(\alpha)\Gamma(\beta)}x^{\alpha-1}(1-x)^{(\beta-1)}
\end{align}
$$

$E(x)=\frac{\alpha}{\alpha+\beta}$ and $Var(x)=\frac{\alpha\beta}{(\alpha+\beta)^2(\alpha+\beta=1)}$
If we reparameterize $\mu=\frac{\alpha}{\alpha+\beta}$ and $\nu=\alpha+\beta$,

$$
f(x) = \frac{\Gamma(\nu)}{\Gamma(\mu\nu)\Gamma(\nu(1-\mu))}x^{\mu\nu-1}(1-x)^{(\nu(1-\mu)-1)}
$$
No Jacobian transformation required since $\mu$ and $\nu$ are parameters of the $f(x)$, not variables.

</details>


<details>
<summary><b>Useful equations (click)</b></summary>

$\psi(\bullet)$ and $\psi_1(\bullet)$ denote for digamma and trigamma functions, 
respectively.

$$
\begin{align}
E[\log{x}] &= \int_{0}^{1}{\log{x}f(x) \,dx} \\
&= \int_{0}^{1}{\log{x} \frac{1}{Beta(\alpha, \beta)}x^{\alpha-1}(1-x)^{(\beta-1)} \,dx} \\
&= \frac{1}{Beta(\alpha, \beta)} \int_{0}^{1}{\log{(x)} x^{\alpha-1}(1-x)^{(\beta-1)} \,dx} \\
&= \frac{1}{Beta(\alpha, \beta)} \int_{0}^{1}{\frac{\partial x^{\alpha-1}(1-x)^{(\beta-1)}}{\partial \alpha} \,dx} \\
&= \frac{1}{Beta(\alpha, \beta)} \frac{\partial}{\partial \alpha}\int_{0}^{1}{ x^{\alpha-1}(1-x)^{(\beta-1)} \,dx} \\
&= \frac{1}{Beta(\alpha, \beta)} \frac{\partial Beta(\alpha, \beta)}{\partial \alpha} \\
&= \frac{\partial \log{[Beta(\alpha, \beta)]}}{\partial \alpha} \\
&= \frac{\partial \log{[\Gamma(\alpha)]}}{\partial \alpha} - \frac{\partial \log{[\Gamma(\alpha + \beta)]}}{\partial \alpha} \\
&= \psi(\alpha) - \psi(\alpha+\beta) 
\end{align}
$$

Similarly, $E[\log{1-x}]=\psi(\beta) - \psi(\alpha+\beta)$.


Furthermore, using $\frac{\partial Beta(\alpha,\beta)}{\partial \alpha} = Beta(\alpha,\beta)\left(\psi(\alpha)-\psi(\alpha+\beta)\right)$ and $\frac{\partial^2 Beta(\alpha,\beta)}{\partial \alpha^2} = Beta(\alpha,\beta)\left(\psi(\alpha)-\psi(\alpha+\beta)\right)^2 + Beta(\alpha,\beta)\left(\psi_1(\alpha)-\psi_1(\alpha+\beta)\right)$,

$$
\begin{align}
E\left[(\log{x})^2\right] &= \frac{1}{Beta(\alpha, \beta)} \frac{\partial^2 Beta(\alpha, \beta)}{\partial \alpha^2} \\
&= \left(\psi(\alpha)-\psi(\alpha+\beta)\right)^2 + \left(\psi_1(\alpha)-\psi_1(\alpha+\beta)\right)
\end{align}
$$

This leads to,

$$
\begin{align}
Var\left[\log{x}\right] &= E\left[(\log{x})^2\right] - E\left[\log{x}\right]^2 \\
&=\psi_1(\alpha)-\psi_1(\alpha+\beta)
\end{align}
$$


</details>




<details>
<summary><b>Continuous IRT (click)</b></summary>

+ Expected item response

$$
\mu = \frac{e^{a(\theta -b)}}{1+e^{a(\theta -b)}} \\
\frac{\partial \mu}{\partial \theta} = a\mu(1-\mu)
$$

+ Probability of a response

$$
f(x)=P(x|\, \theta, a, b, \nu) = \frac{\Gamma(\nu)}{\Gamma(\mu\nu)\Gamma(\nu(1-\mu))} x^{\mu\nu-1} (1-x)^{\nu(1-\mu)-1} \\
$$

+ Gradient

$$
\log{f} = \log{\Gamma(\nu)}-\log{\Gamma(\mu\nu)}-\log{\Gamma(\nu(1-\mu))} + (\mu\nu-1)\log{x} + (\nu(1-\mu)-1) \log{(1-x)}
$$

$$
\frac{\partial \log{f}}{\partial \theta} = a\nu\mu(1-\mu)\left[-\psi{(\mu\nu)}+\psi{(\nu(1-\mu))}+ \log{\left(\frac{x}{1-x}\right)}\right]
$$

+ Information


$$
E\left[ \left(\frac{\partial \log{f}}{\partial \theta}\right)^2 \right] = 
(a\nu\mu(1-\mu))^2\left[E\left[ \log{\left(\frac{x}{1-x}\right)^2}\right] 
-2 \left(\psi{(\mu\nu)}-\psi{(\nu(1-\mu))}\right )E\left[ \log{\left(\frac{x}{1-x}\right)}\right]
+\left(\psi{(\mu\nu)}-\psi{(\nu(1-\mu))}\right )^2
\right] 
$$

$$
\begin{align}
E\left[ \log{\left(\frac{x}{1-x}\right)^2}\right] 
&= E\left[ \log{\left(x\right)^2}\right]
-2 E\left[ \log{\left(x\right)}\log{\left(1-x\right)}\right]
+ E\left[ \log{\left(1-x\right)^2}\right] \\
&= Var\left[ \log{\left(x\right)}\right]+E\left[ \log{\left(x\right)}\right]^2 \\
&\qquad -2 Cov\left[ \log{\left(x\right)}\log{\left(1-x\right)}\right]-2E\left[ \log{\left(x\right)}\right]E\left[ \log{\left(1-x\right)}\right] \\
&\qquad + Var\left[ \log{\left(1-x\right)}\right]+E\left[ \log{\left(1-x\right)}\right]^2 \\
&= \psi_{1}(\alpha)-\psi_{1}(\alpha+\beta) +E\left[ \log{\left(x\right)}\right]^2 \\
&\qquad +2 \psi_{1}(\alpha+\beta)-2E\left[ \log{\left(x\right)}\right]E\left[ \log{\left(1-x\right)}\right] \\
&\qquad + \psi_{1}(\beta)-\psi_{1}(\alpha+\beta)+E\left[ \log{\left(1-x\right)}\right]^2 \\
&= \psi_{1}(\alpha)-\psi_{1}(\alpha+\beta) +\left[ \psi(\alpha)-\psi(\alpha+\beta)\right]^2 \\
&\qquad +2 \psi_{1}(\alpha+\beta)-2 \left(\psi(\alpha)-\psi(\alpha+\beta)\right)\left(\psi(\beta)-\psi(\alpha+\beta)\right)  \\
&\qquad + \psi_{1}(\beta)-\psi_{1}(\alpha+\beta)+\left[\psi(\beta)-\psi(\alpha+\beta)\right]^2 \\
&= \psi_{1}(\alpha) +\psi_{1}(\beta) +\left[\psi(\alpha)-\psi(\beta)\right]^2
\end{align}
$$

$$
\begin{align}
E\left[\left(\frac{\partial \log{f}}{\partial \theta}\right)^2 \right] & = 
(a\nu\mu(1-\mu))^2\left[E\left[ \log{\left(\frac{x}{1-x}\right)^2}\right] 
-2 \left(\psi{(\mu\nu)}-\psi{(\nu(1-\mu))}\right )E\left[ \log{\left(\frac{x}{1-x}\right)}\right]
+\left(\psi{(\mu\nu)}-\psi{(\nu(1-\mu))}\right )^2
\right] \\
&= (a\nu\mu(1-\mu))^2\left[\psi_{1}(\alpha) +\psi_{1}(\beta)  
+\left[\psi(\alpha)-\psi(\beta)\right]^2
-2 \left(\psi{(\alpha)}-\psi{(\beta)}\right )\left(\psi{(\alpha)}-\psi{(\beta)}\right )
+\left(\psi{(\alpha)}-\psi{(\beta)}\right )^2
\right] \\
&= (a\nu\mu(1-\mu))^2\left[\psi_{1}(\alpha) +\psi_{1}(\beta)  
\right] \\
\end{align}
$$

$$
I(\theta) = E\left[\left(\frac{\partial \log{f}}{\partial \theta}\right)^2 \right] =  (a\nu\mu(1-\mu))^2\left[\psi_{1}(\alpha) +\psi_{1}(\beta)\right]
$$



</details>

\
\

+ **Preparing data**

The function `DataGeneration` can be used in a preparation step.
This function returns a set of artificial data and the true parameters underlying the data.

```{r}
Alldata <- DataGeneration(N=500,
                          nitem_C = 4,
                          latent_dist = "2NM",
                          a_l = .3, 
                          a_u = .7,
                          d = 1.664,
                          sd_ratio = 2,
                          prob = 0.3)

data <- Alldata$data_C
theta <- Alldata$theta
colnames(data) <- paste0("item", 1:4)
```

\

+ **Analysis** (parameter estimation)

```{r, results='hide', message=FALSE}
Mod1 <- IRTest_Cont(data = data,
                    latent_dist = "2NM",
                    range = c(-4,4),
                    q = 81,
                    threshold = 0.01)
```

\

+ **Results**

```{r, message=FALSE, fig.align='center', fig.height=6, fig.width=6}
### Summary
summary(Mod1)

### Log-likelihood
logLik(Mod1)

### The estimated item parameters
coef(Mod1)

### The estimated ability parameters
fscore <- factor_score(Mod1, ability_method = "MLE")
plot(theta, fscore$theta)
abline(b=1, a=0)

### Standard errors of ability parameter estimates
plot(fscore$theta, fscore$theta_se)
```


* The result of latent distribution estimation

```{r, fig.align='center', fig.asp=.6, fig.width=6}
plot(Mod1, mapping = aes(colour="Estimated"), linewidth = 1) +
  lims(y = c(0, .75))+
  geom_line(
    mapping=aes(
      x=seq(-6,6,length=121), 
      y=dist2(seq(-6,6,length=121), prob = .3, d = 1.664, sd_ratio = 2), 
      colour="True"),
    linewidth = 1)+
  labs(title="The estimated latent density using '2NM'", colour= "Type")+
  theme_bw()
```

* Item response function

```{r, fig.align='center', fig.asp=0.8, fig.width=6}
p1 <- plot_item(Mod1,1)
p2 <- plot_item(Mod1,2)
p3 <- plot_item(Mod1,3)
p4 <- plot_item(Mod1,4)
grid.arrange(p1, p2, p3, p4, ncol=2, nrow=2)
```


* Reliability

```{r}
reliability(Mod1)
```

* Posterior distributions for the examinees

Each examinee's posterior distribution is identified in the E-step of the estimation algorithm (i.e., EM algorithm).
Posterior distributions can be found in `Mod1$Pk`.

```{r, fig.align='center', fig.asp=.7, fig.width=6}
set.seed(1)
selected_examinees <- sample(1:500,6)
post_sample <- 
  data.frame(
    X = rep(seq(-4,4, length.out=81),6), 
    prior = rep(Mod1$Ak/(Mod1$quad[2]-Mod1$quad[1]), 6),
    posterior = 10*c(t(Mod1$Pk[selected_examinees,])), 
    ID = rep(paste("examinee", selected_examinees), each=81)
    )

ggplot(data=post_sample, mapping=aes(x=X))+
  geom_line(mapping=aes(y=posterior, group=ID, color='Posterior'))+
  geom_line(mapping=aes(y=prior, group=ID, color='Prior'))+
  labs(title="Posterior densities for selected examinees", x=expression(theta), y='density')+
  facet_wrap(~ID, ncol=2)+
  theme_bw()
```

* Test information function

```{r, fig.align='center', fig.asp=.8, fig.width=6}
ggplot()+
  stat_function(
    fun = inform_f_test,
    args = list(Mod1)
  )+ 
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 1),
    mapping = aes(color="Item 1")
  )+
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 2),
    mapping = aes(color="Item 2")
  )+
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 3),
    mapping = aes(color="Item 3")
  )+
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 4),
    mapping = aes(color="Item 4")
  )+
  lims(x=c(-6,6))+
  labs(title="Test information function", x=expression(theta), y='information')+
  theme_bw()
```


----------

\break

# 4. Mixed-format test

+ **Preparing data**

As in the cases of dichotomous and polytomous items, the function `DataGeneration` can be used in the preparation step.
This function returns artificial data and some useful objects for analysis (i.e., `theta`, `data_D`, `item_D`, `data_P`, & `item_P`).


```{r}
Alldata <- DataGeneration(model_D = 2,
                          model_P = "GRM",
                          N=1000,
                          nitem_D = 10,
                          nitem_P = 5,
                          latent_dist = "2NM",
                          d = 1.664,
                          sd_ratio = 1,
                          prob = 0.5)

DataD <- Alldata$data_D
DataP <- Alldata$data_P
theta <- Alldata$theta
colnames(DataD) <- paste0("item", 1:10)
colnames(DataP) <- paste0("item", 1:5)
```


\
\
\

+ **Analysis** (parameter estimation)

```{r, results='hide', message=FALSE}
Mod1 <- IRTest_Mix(data_D = DataD,
                   data_P = DataP,
                   model_D = "2PL",
                   model_P = "GRM",
                   latent_dist = "KDE")
```

\
\
\

+ **Results**

```{r, message=FALSE, fig.align='center', fig.height=6, fig.width=6}
### Summary
summary(Mod1)

### Log-likelihood
logLik(Mod1)

### The estimated item parameters
coef(Mod1)

### Standard errors of the item parameter estimates
coef_se(Mod1)

### The estimated ability parameters
fscore <- factor_score(Mod1, ability_method = "MLE")
plot(theta, fscore$theta)
abline(b=1, a=0)

### Standard errors of ability parameter estimates
plot(fscore$theta, fscore$theta_se)
```

* The result of latent distribution estimation

```{r, fig.align='center', fig.asp=.6, fig.width=6}
plot(Mod1, mapping = aes(colour="Estimated"), linewidth = 1) +
  stat_function(
    fun = dist2,
    args = list(prob = .5, d = 1.664, sd_ratio = 1),
    mapping = aes(colour = "True"),
    linewidth = 1) +
  lims(y = c(0, .75)) + 
  labs(title="The estimated latent density using '2NM'", colour= "Type")+
  theme_bw()
```


* Item response function

```{r, fig.align='center', fig.asp=0.8, fig.width=6}
p1 <- plot_item(Mod1,1, type="d")
p2 <- plot_item(Mod1,4, type="d")
p3 <- plot_item(Mod1,8, type="d")
p4 <- plot_item(Mod1,10, type="d")
grid.arrange(p1, p2, p3, p4, ncol=2, nrow=2)
```

```{r, fig.align='center', fig.asp=0.8, fig.width=6}
p1 <- plot_item(Mod1,1, type="p")
p2 <- plot_item(Mod1,2, type="p")
p3 <- plot_item(Mod1,3, type="p")
p4 <- plot_item(Mod1,4, type="p")
grid.arrange(p1, p2, p3, p4, ncol=2, nrow=2)
```



+ **Item fit**
```{r}
item_fit(Mod1)
```

* Reliability

```{r}
reliability(Mod1)
```

* Posterior distributions for the examinees

Each examinee's posterior distribution is identified in the E-step of the estimation algorithm (i.e., EM algorithm).
Posterior distributions can be found in `Mod1$Pk`.

```{r, fig.align='center', fig.asp=.7, fig.width=6}
set.seed(1)
selected_examinees <- sample(1:1000,6)
post_sample <- 
  data.frame(
    X = rep(seq(-6,6, length.out=121),6), 
    prior = rep(Mod1$Ak/(Mod1$quad[2]-Mod1$quad[1]), 6),
    posterior = 10*c(t(Mod1$Pk[selected_examinees,])), 
    ID = rep(paste("examinee", selected_examinees), each=121)
    )

ggplot(data=post_sample, mapping=aes(x=X))+
  geom_line(mapping=aes(y=posterior, group=ID, color='Posterior'))+
  geom_line(mapping=aes(y=prior, group=ID, color='Prior'))+
  labs(title="Posterior densities for selected examinees", x=expression(theta), y='density')+
  facet_wrap(~ID, ncol=2)+
  theme_bw()
```

* Test information function

```{r, fig.align='center', fig.asp=.8, fig.width=8}
ggplot()+
  stat_function(
    fun = inform_f_test,
    args = list(Mod1)
  )+ 
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 1, "d"),
    mapping = aes(color="Dichotomous Item 1")
  )+
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 2, "d"),
    mapping = aes(color="Dichotomous Item 2")
  )+
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 3, "d"),
    mapping = aes(color="Dichotomous Item 3")
  )+
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 1, "p"),
    mapping = aes(color="Polytomous Item 1")
  )+
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 2, "p"),
    mapping = aes(color="Polytomous Item 2")
  )+
  stat_function(
    fun=inform_f_item,
    args = list(Mod1, 3, "p"),
    mapping = aes(color="Polytomous Item 3")
  )+
  lims(x=c(-6,6))+
  labs(title="Test information function", x=expression(theta), y='information')+
  theme_bw()
```


----------

\break

# 5. Model comparison

+ **Data generation and model fitting**

```{r}
data <- DataGeneration(N=1000,
                       nitem_D = 10,
                       latent_dist = "2NM",
                       d = 1.664,
                       sd_ratio = 2,
                       prob = 0.3)$data_D
```

```{r, results='hide', message=FALSE}
model_fits <- list()
model_fits[[1]] <- IRTest_Dich(data)
model_fits[[2]] <- IRTest_Dich(data, latent_dist = "EHM")
model_fits[[3]] <- IRTest_Dich(data, latent_dist = "2NM")
model_fits[[4]] <- IRTest_Dich(data, latent_dist = "KDE")
for(i in 1:10){
  model_fits[[i+4]] <- IRTest_Dich(data, latent_dist = "DC", h = i)
}

names(model_fits) <- c("Normal", "EHM", "2NM", "KDM", paste0("DC", 1:10))
```

+ **The best Davidian-curve model**

```{r}
do.call(what = "anova", args = model_fits[5:14])
do.call(what = "best_model", args = model_fits[5:14])
```

+ **The best model overall**

```{r}
do.call(what = "best_model", args = c(model_fits[c(1:4,5)], criterion ="AIC"))
```






