---
title: "IRT without the normality assumption"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{IRT without the normality assumption}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
fontsize: 12pt
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(IRTest)
library(ggplot2)
```

\
\
\

# 1. Dichotomous items

+ **Preparing data**

The function `DataGeneration` can be used for the pre-analysis step.
This function returns artificial data and some useful objects for analysis (i.e., `theta`, `data_D`, `item_D`, and `initialitem_d`).
\
\
In the parameter estimation process, 
the `initialitem` can be used for an input of the function `IRTest_Dich` (i.e., `initialitem = initialitem`).
The `data` is an artificial item response data that could be unnecessary if user-imported item response data is used.
The `theta` and `item` are not used for the estimation process.
They can be considered as the true parameters only if the artificial data (`data`) is used for analysis.

```{r}
Alldata <- DataGeneration(seed = 123456789,
                          model_D = rep(1:2, each=5),
                          N=1000,
                          nitem_D = 10,
                          nitem_P = 0,
                          d = 1.664,
                          sd_ratio = 2,
                          prob = 0.3)

data <- Alldata$data_D
item <- Alldata$item_D
initialitem <- Alldata$initialitem_D
theta <- Alldata$theta
```

If the artificial data (`data`) is used, the true latent distribution looks like,

```{r, fig.align='center', fig.height=2.5, fig.width=6, echo=F}
ggplot(data = data.frame(x = c(-6,6)), aes(x)) +
  stat_function(fun = dist2, n = 101, args = list(prob = 0.3, d=1.664, sd_ratio = 2)) +
  ylab("latent density") +
  xlab(expression(theta)) +
  scale_y_continuous(breaks = NULL)
```

\
\
\

+ **Analysis** (parameter estimation)

```{r, results='hide', message=FALSE}
######                            ######
###### Empirical histogram method ######
######                            ######
Mod1 <- IRTest_Dich(initialitem = initialitem,
                    data = data,
                    model = rep(1:2, each=5),
                    latent_dist = "EHM",
                    max_iter = 200,
                    threshold = .0001)

######                                  ######
###### Kernel density estimation method ######
######                                  ######
# Mod1 <- IRTest_Dich(initialitem = initialitem,
#                     data = data,
#                     model = rep(1:2, each=5),
#                     latent_dist = "KDE",
#                     bandwidth = "SJ-ste",
#                     max_iter = 200,
#                     threshold = .001)

######                      ######
###### Normality assumption ######
######                      ######
#  Mod1 <- IRTest_Dich(initialitem = initialitem,
#                      data = data,
#                      model = rep(1:2, each=5),
#                      latent_dist = "Normal",
#                      max_iter = 200,
#                      threshold = .0001)

######                                             ######
###### Two-component Gaussian mixture distribution ######
######                                             ######
#  Mod1 <- IRTest_Dich(initialitem = initialitem,
#                      data = data,
#                      model = rep(1:2, each=5),
#                      latent_dist = "Mixture",
#                      max_iter = 200,
#                      threshold = .0001)

######                                                       ######
###### Davidian curve (for an arbitrarily chosen case of h=4)######
######                                                       ######
#  Mod1 <- IRTest_Dich(initialitem = initialitem,
#                      data = data,
#                      model = rep(1:2, each=5),
#                      latent_dist = "DC",
#                      max_iter = 200,
#                      threshold = .0001,
#                      h=4)
```

\
\
\

+ **Results**

```{r, fig.align='center', fig.height=2.5, fig.width=6}
### The estimated item parameters
Mod1$par_est

### The asymptotic standard errors of item parameters
Mod1$se

### The estimated ability parameters
head(Mod1$theta)

### The estimated latent distribution
plot_LD(Mod1, xlim = c(-6,6))
```

\
\

----------

\break

# 2. Polytomous items

+ **Preparing data**

As in the case of dichotomous items, the function `DataGeneration` can be used for the pre-analysis step.
This function returns artificial data and some useful objects for analysis (i.e., `theta`, `data_P`, `item_P`, and `initialitem_P`).
\
\
In the parameter estimation process, 
the `initialitem` can be used for an input of the function `IRTest_Poly` (i.e., `initialitem = initialitem`).
The `data` is an artificial item response data that could be unnecessary if user-imported item response data is used.
The `theta` and `item` are not used for the estimation process.
They can be considered as the true parameters only if the artificial data (`data`) is used for analysis.


```{r}
Alldata <- DataGeneration(seed = 123456789,
                          model_P = "GPCM",
                          categ = rep(c(3,7), each = 5),
                          N=1000,
                          nitem_D = 0,
                          nitem_P = 10,
                          d = 1.414,
                          sd_ratio = 2,
                          prob = 0.5)

data <- Alldata$data_P
item <- Alldata$item_P
initialitem <- Alldata$initialitem_P
theta <- Alldata$theta
```

If the artificial data (`data`) is used, the true latent distribution looks like,

```{r, fig.align='center', fig.height=2.5, fig.width=6, echo=F}
ggplot(data = data.frame(x = c(-6,6)), aes(x)) +
  stat_function(fun = dist2, n = 101, args = list(prob = 0.5, d=1.414, sd_ratio = 2)) +
  ylab("latent density") +
  xlab(expression(theta)) +
  scale_y_continuous(breaks = NULL)
```

\
\
\

+ **Analysis** (parameter estimation)

```{r, results='hide', message=FALSE}
######                                  ######
###### Kernel density estimation method ######
######                                  ######
Mod1 <- IRTest_Poly(initialitem = initialitem,
                    data = data,
                    model = "GPCM",
                    latent_dist = "KDE",
                    bandwidth = "SJ-ste",
                    max_iter = 200,
                    threshold = .001)

######                      ######
###### Normality assumption ######
######                      ######
#  Mod1 <- IRTest_Poly(initialitem = initialitem,
#                      data = data,
#                      model = "GPCM",
#                      latent_dist = "Normal",
#                      max_iter = 200,
#                      threshold = .001)

######                            ######
###### Empirical histogram method ######
######                            ######
#  Mod1 <- IRTest_Poly(initialitem = initialitem,
#                      data = data,
#                      model = "GPCM",
#                      latent_dist = "EHM",
#                      max_iter = 200,
#                      threshold = .001)

######                                             ######
###### Two-component Gaussian mixture distribution ######
######                                             ######
#  Mod1 <- IRTest_Poly(initialitem = initialitem,
#                      data = data,
#                      model = "GPCM",
#                      latent_dist = "Mixture",
#                      max_iter = 200,
#                      threshold = .001)

######                                                        ######
###### Davidian curve (for an arbitrarily chosen case of h=4) ######
######                                                        ######
#  Mod1 <- IRTest_Poly(initialitem = initialitem,
#                      data = data,
#                      model = "GPCM",
#                      latent_dist = "DC",
#                      max_iter = 200,
#                      threshold = .001,
#                      h=4)
```

\
\
\

+ **Results**

```{r, fig.align='center', fig.height=2.5, fig.width=6}
### The estimated item parameters
Mod1$par_est

### The asymptotic standard errors of item parameters
Mod1$se

### The estimated ability parameters
head(Mod1$theta)

### The estimated latent distribution
plot_LD(Mod1, xlim = c(-6,6))
```

\
\

----------

\break

# 3. Mixed-format test

+ **Preparing data**

As in the case of dichotomous and polytomous items, the function `DataGeneration` can be used for the pre-analysis step.
This function returns artificial data and some useful objects for analysis (i.e., `theta`, `data_D`, `item_D`, `initialitem_D`, `data_P`, `item_P`, and `initialitem_P`).
\
\
In the parameter estimation process, 
the `initialitem` can be used for an input of the function `IRTest_Mix` (i.e., `initialitem = initialitem`).
The `data` is an artificial item response data that could be unnecessary if user-imported item response data is used.
The `theta` and `item` are not used for the estimation process.
They can be considered as the true parameters only if the artificial data (`data`) is used for analysis.


```{r}
Alldata <- DataGeneration(seed = 123456789,
                          model_D = rep(2,5),
                          model_P = "GPCM",
                          categ = rep(3,5),
                          N=1000,
                          nitem_D = 5,
                          nitem_P = 5,
                          d = 1.664,
                          sd_ratio = 1,
                          prob = 0.5)

DataD <- Alldata$data_D
DataP <- Alldata$data_P
itemD <- Alldata$item_D
itemP <- Alldata$item_P
initialitemD <- Alldata$initialitem_D
initialitemP <- Alldata$initialitem_P
theta <- Alldata$theta
```

If the artificial data (`data`) is used, the true latent distribution looks like,

```{r, fig.align='center', fig.height=2.5, fig.width=6, echo=F}
ggplot(data = data.frame(x = c(-6,6)), aes(x)) +
  stat_function(fun = dist2, n = 101, args = list(prob = 0.5, d=1.664, sd_ratio = 1)) +
  ylab("latent density") +
  xlab(expression(theta)) +
  scale_y_continuous(breaks = NULL)
```

\
\
\

+ **Analysis** (parameter estimation)

```{r, results='hide', message=FALSE}
######                                  ######
###### Kernel density estimation method ######
######                                  ######
Mod1 <- IRTest_Mix(initialitem_D = initialitemD,
                   initialitem_P = initialitemP,
                   data_D = DataD,
                   data_P = DataP,
                   model_D = rep(2,5),
                   model_P = "GPCM",
                   latent_dist = "KDE",
                   bandwidth = "SJ-ste",
                   max_iter = 200,
                   threshold = .001)

######                      ######
###### Normality assumption ######
######                      ######
#  Mod1 <- IRTest_Mix(initialitem_D = initialitemD,
#                     initialitem_P = initialitemP,
#                     data_D = DataD,
#                     data_P = DataP,
#                     model_D = rep(2,5),
#                     model_P = "GPCM",
#                     latent_dist = "Normal",
#                     max_iter = 200,
#                     threshold = .001)

######                            ######
###### Empirical histogram method ######
######                            ######
#  Mod1 <- IRTest_Mix(initialitem_D = initialitemD,
#                     initialitem_P = initialitemP,
#                     data_D = DataD,
#                     data_P = DataP,
#                     model_D = rep(2,5),
#                     model_P = "GPCM",
#                     latent_dist = "EHM",
#                     max_iter = 200,
#                     threshold = .001)

######                                             ######
###### Two-component Gaussian mixture distribution ######
######                                             ######
#  Mod1 <- IRTest_Mix(initialitem_D = initialitemD,
#                     initialitem_P = initialitemP,
#                     data_D = DataD,
#                     data_P = DataP,
#                     model_D = rep(2,5),
#                     model_P = "GPCM",
#                     latent_dist = "Mixture",
#                     max_iter = 200,
#                     threshold = .001)

######                                                        ######
###### Davidian curve (for an arbitrarily chosen case of h=4) ######
######                                                        ######
#  Mod1 <- IRTest_Mix(initialitem_D = initialitemD,
#                     initialitem_P = initialitemP,
#                     data_D = DataD,
#                     data_P = DataP,
#                     model_D = rep(2,5),
#                     model_P = "GPCM",
#                     latent_dist = "DC",
#                     max_iter = 200,
#                     threshold = .001,
#                     h = 4)
```

\
\
\

+ **Results**

```{r, fig.align='center', fig.height=2.5, fig.width=6}
### The estimated item parameters
Mod1$par_est

### The asymptotic standard errors of item parameters
Mod1$se

### The estimated ability parameters
head(Mod1$theta)

### The estimated latent distribution
plot_LD(Mod1, xlim = c(-6,6))
```

\
\
----

